---
title: "Bootstrap"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## An review of Sampling Distributions (how and why):

1. Start with a *known* population. That is, I know something about its distribution.
2. Generate a *sampling distribution* for some statistic (mean, max, proportion, etc.):
	a. Pull a random sample
	b. Compute a statistic (mean, max, etc.) of that sample.
	c. Record that value
	d. Repeat
	e. The collection of these sample statistics *is* the sample distribution.
8. Use the sampling distribution to compute probabilities:
	* Eg. What is the probability that the mean of a sample is ____?

## The problem: We *don’t know* the population.

## The *bootstrap* solution: 

* Use a sample *in place of* the population, 
* generate a *bootstrap* distribution from this sample (we’ll get there)
* Ask: Can I use the bootstrap distribution in the same way I’d use a (theoretical) sampling distribution?
* If not, what can I do with it?




## How to contruct a *bootstrap* distribution and comparison to sampling distribution

In order to do the comparison part we'll start with a *known* distribution.  

Keep in mind this is the part we can't do in practice.  We do it here to see if a bootstrap distribution is of any value.

### Known distribution:

normal with mean $\mu=36$ and standard deviation $\sigma = 8$.

Here is the distribution:

```{r}
library(ggfortify)
mu <- 36
sigma <- 8
ggdistribution(dnorm, seq(0, 72, 0.1), mean=mu, sd=sigma)
```

We know from the CLT that the sampling distribution of the mean for samples of size $n=200$ is *normal*, has mean $\mu=36$, and standard error $\text{S.E.}=\frac{\sigma}{\sqrt{n}}$.  Here is its plot:

```{r}
n <- 200
ggdistribution(dnorm, seq(0, 72, 0.1), mean=mu, sd=sigma/sqrt(n))
```


This is the *theoretical* sampling distribution. Theoretical in the sense that we started with a *known* population.

### Bootstrap method

* We'll start by drawing a random sample of size $n=200$ (same size as in the sampling distribution) from the population:

```{r}
BaseSample <- rnorm(n, mean=mu, sd=sigma)
```

Look at it so you know what it is:

```{r}
head(BaseSample)
```

It is a vector of number chosen randomly from the population distribution.

* Now, in the bootstrap procedure we are going to treat `BaseSample` as the population.  From here until the end pretend we *don't know* the population.

* We are going to resample and compute the mean of each resample, just as we would to find a sampling distribution, but:

  * We are going to pick resamples of size 200!  The same size as `BaseSample`.
  * This only makes sense if we allow replacement: `replace=T`.  Otherwise we simply get the same sample over and over.
  
  
  
* We repeat, say 1000 times.
* the result is the *bootstrap distribution*:

```{r}
N <- 1000
BootDist <- numeric(N)
for (i in 1:N)
{samp <- sample(BaseSample, n, replace=T)
BootDist[i] <- mean(samp)}
```

Look at it to see what you've created:

```{r}
head(BootDist)
```

This is a vector whose entries are means of each of the resamples. There are 1000 entries in this vector.

Look at the graphical representation of this distribution:

```{r}
hist(BootDist)
```

We'll plot the *sampling distribution* on a similar scale:

```{r}
ggdistribution(dnorm, seq(34, 40, 0.1), mean=mu, sd=sigma/sqrt(n))
```

* These two distributions:

  * have a similar shape - namely normal
  * have a similar spread - most of the values are within a range of about 4
  * do not have the same center: the theoretical sampling distribution is centered at 36, the bootstrap distribution is centered, maybe, at 36.5

* Let's do this precisely:

  * We *know* the mean of the sampling distribution is 36
  * We *know* the standard deviation of the sampling distribution is:
```{r}
sigma/sqrt(n)
```
  
  * The mean of the bootstrap distribution is:
```{r}
mean(BootDist)
```

(depending on my knit this value may be close to 36, but not always)
  
  * The standard deviation of the bootstrap distribution is:
```{r}
sd(BootDist)
```

One more comparison:

```{r}
mean(BaseSample)
```

The HW asks for a table.  Here is quick and dirty way to make a table:

```{r}
Distribution <- c("Population", "Sample Distribution", "Sample", "Bootstrap Distribution")
means <- c(mu, mu, mean(BaseSample), mean(BootDist))
sdevs <- c(sigma, sigma/sqrt(n), sd(BaseSample),sd(BootDist))
results <- cbind(Distribution, means, sdevs)
knitr::kable(results, digits=2)
```


So, the mean of the bootstrap distribuion is similar to the mean of `BaseSample`.  This is not unexpected, it also isn't particularly useful.  We want to say something about the mean of *real* popultion, not the mean of the sample.

We'll keep this in mind: 

* Bootstraping does a good job giving the *shape* of the sampling distribution.
* Bootstraping does a good job giving the *spread* of the sampling distribution.
* Bootstraping does *not* do a good job giving the *center* of the sampling distribuiton.

We'll see how to use this next....