---
title: "Exam_One_A"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





# Terminology
1. What is the difference between a *population* and a *sample*?

Answer: 


2. What is the difference between a *statistic* and a *parameter*.  Give an example of a *statistic*.

Answer:


3. In a data set, what is the difference between an *observation* and a *variable*?

Answer:



# Exploratory Data Analysis

Use one of the following two chunks to load the data set `Alelager`.

If you have the `resampledata` package installed use:

```{r}
library(resampledata)
BeerData<-Alelager
head(BeerData)
```

**or**, you can use:

```{r}
BeerData<-read.csv("https://sites.google.com/site/chiharahesterberg/data2/Alelager.csv")
head(BeerData)
```

1. Create side-by-side box plots of the alcohol content, grouped by beer type. Comment on the information the box plots convey.

```{r}

```

Commentary: 


2. Create a histogram of the calorie content of all beers in the data set. Describe the distribution.
```{r}

```

Description of distribution:

3.  Sampling Distribution
*	Simulate the sampling distribution of the mean of the number of calories, for samples of size 3.
		
```{r}

```



*	Create a histogram and describe the simulated sampling distribution.
		
```{r}

```


Description of sampling distribution:		
		
*	Use the simulation to find the probability that the average calorie content of 3 beers is less than 150 calories.
		
```{r}
```
		
* Use the simulation to find the probability that the average calorie content of 3 beers is more than 160 calories.
  
```{r}
```
  
  
4. The following code generates two ecdfs, one for calories of ales, one for calories of lagers. 
 
```{r}
Ales <- subset(BeerData, select=Calories, subset=Type=="Ale", drop=T)
Lagers <- subset(BeerData, select=Calories, subset=Type=="Lager", drop=T)
#plot.new()
plot.ecdf(Ales, col="red")
plot.ecdf(Lagers, col="blue", add=TRUE)
abline(v=165)
```

What percentage of ales in the data set have less than 165 calories? 

What percentage of lagers in the data set have less than 165 calories?


# Permutation test

Conduct a permutation test to determine whether there is a difference in the mean calorie content between ales and lagers. Your test should start with a statement of the formal hypotheses and finish with a conclusion.

Put each step in its own chunk and **clearly explain what each chunk is doing and what information the output provides.**

```{r}

```




```{r}

```



```{r}

```


```{r}

```



```{r}

```


(I put in several r-chunks.  You might use more, you might use fewer.)


# Critical reading

Below, I conduct a permutation test to test whether there was, on average, a difference in the annual snow fall in Colorado in the years 2019 and 2020.  Identify the mistake and fix it.

First, we load and briefly inspect the data:
```{r}
SnowData <- read.csv("https://www.dropbox.com/s/h8lkjdqoqui3bzw/TestDownload2.csv?dl=1")
head(SnowData)
```

The data shows annual snow fall recorded at 51 NOAA weather station in Colorado in 2019, and 2020. 


The formal statements of the hypotheses are:

$$H_0: \mu_{d_{2020}-d_{2019}}=0$$

$$H_A: \mu_{d_{2020}-d_{2019}}\neq 0$$

where $d_{2020}-d_{2019}$ represents the difference in snow fall by station. 

We'll compute the difference in snowfall at each station and compute the observed mean of differences:

```{r}
DiffAtStations<-SnowData$Snow2020-SnowData$Snow2019
Observed<-mean(DiffAtStations)
Observed
```

Next, we'll pool the data in order to find a permutation distribution:

```{r}
PooledData<- c(SnowData$Snow2020, SnowData$Snow2019)
```

Here is the loop that generates a permutation distribution:


```{r}
N=10^4
SampleDist<-numeric(N)
for (i in 1:N)
{index<-sample(102,51, replace=F)
SampleDist[i]<- mean(PooledData[index]-PooledData[-index])}
```



Finally, we compute the $p$-value for our observed statistic:

```{r}
2*min((sum(SampleDist >= Observed)+1)/(N+1),
(sum(SampleDist <= Observed)+1)/(N+1))
```

Since the $p$-value is small we reject the null hypothesis and conclude that the mean difference in snow depth between 2019 and 2020 was different.



**Your notes are below.  Do NOT delete the following chunk or your file may not knit.
```{r}
knitr::knit_exit()
```


*Exploratory statistics*

normal dist = qnorm

ecdf:
plot.new()
plot.ecdf(smoker, col="red")
plot.ecdf(nonsmoker, col="blue", add=TRUE)

qqplots:
plot.new()
qqnorm(Spruce$Height5)
qqline(Spruce$Height5)

barplot(table(FlightData$Carrier))

Contingency tables:
table(FlightData$Carrier, FlightData$Delayed30)
prop.table(table(FlightData$Carrier, FlightData$Delayed30))
prop.table(table(FlightData$Carrier, FlightData$Delayed30),1)
prop.table(table(FlightData$Carrier, FlightData$Delayed30),2)

tapply(subset, subset, mean or sd)
boxplot(subset~subset) *remove<-subset(FlightData,subset=Delay<=100)boxplot(remove$Delay~remove$Destination)*

quantile(subset)

*Permutation Tests*

$$H_0: \mu_{UA}-\mu_{AA}=0.$$
$$H_A: \mu_{UA}-\mu_{AA}\neq 0.$$

subset(title, select=certain area, subset="titleofcertainportionwithinsubset"=="UA" , drop=T)

observed is mean(subset)-mean(othersubset), propone-propother

index <- sample(totalsizeofpop,size=portionofpop, replace=F)

**resampling**
N <- 10^4-1
result <- numeric(N)
for (i in 1:N)
{index <- sample(4029,size=1123, replace=F)
result[i] <- mean(alldelays[index])-mean(alldelays[-index])
} 

alldelays <- delays$Delay
poolsize <- length(alldelays)
poolsize
GroupOneSize <- length(Maydelays)
GroupOneSize
GroupTwoSize <- length(Junedelays)
GroupTwoSize

N <- 10^4-1
result <- numeric(N)
for (i in 1:N)
{index <- sample(poolsize,size=GroupOneSize, replace=F)
result[i] <- (sum(alldelays[index]>20)/GroupOneSize)-(sum(alldelays[-index]>20)/GroupTwoSize)
} 

hist(result)

pvalue <- 2*min((sum(result >= observed)+1)/(N+1),
(sum(result <= observed)+1)/(N+1)) *2times the small tail
 *compare to <.05 proves null hypothesis*
 
*Sampling Distributions*
head()
hist()

2 pop:
N=10^4
Xsumofmax <- numeric(N)
for (i in 1:N)
{sampleA <- sample(popA, size=3, replace=F)
sampleB <- sample(popB, size=3, replace=F)
Xsumofmax[i] <- max(sampleA)+max(sampleB)}

 1 pop:
 N=10^4
Xbar <- numeric(N)
for (i in 1:N)
  {mysample <- sample(ChiMarathonMen$FinishMin, size=6,replace=F)
  Xbar[i] <- mean(mysample)}

hist(result)
mean(result)
sd(reslut) *compare to mean and sd of topic*
sderror is sd/sqrt(samplesize)

*random extra*

library()
read('url')
boxplot(subset~subset, drop=T)
length(title)
abline(v=observed, col="blue")
proportion = sum(one<>#)/length(other)
stripchart(title, method="stack", at=0)
skewness()
kurtosis()
*hist(), change widths* min(UAOnly$Delay) max(UAOnly$Delay) width=seq(-50,700, by=25)
hist(UAOnly$Delay, width)
hist(AAOnly$Delay, width)
para:chpop
stat:numchdata
cens:offcouorsurvpop
statin:concpopbassam
samf:rechsam
undcov:ompepfrpop
tstst:numfundatadetrestst
stsam:popdivstraa,samfrsttm






