---
title: "Exam_One_A"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





# Terminology
1. What is the difference between a *population* and a *sample*?

Answer: 


2. What is the difference between a *statistic* and a *parameter*.  Give an example of a *statistic*.

Answer:


3. In a data set, what is the difference between an *observation* and a *variable*?

Answer:



# Exploratory Data Analysis

Use one of the following two chunks to load the data set `Alelager`.

If you have the `resampledata` package installed use:

```{r}
library(resampledata)
BeerData<-Alelager
head(BeerData)
```

**or**, you can use:

```{r}
BeerData<-read.csv("https://sites.google.com/site/chiharahesterberg/data2/Alelager.csv")
head(BeerData)
```

1. Create side-by-side box plots of the alcohol content, grouped by beer type. Comment on the information the box plots convey.

```{r}

```

Commentary: 


2. Create a histogram of the calorie content of all beers in the data set. Describe the distribution.
```{r}

```

Description of distribution:

3.  Sampling Distribution
*	Simulate the sampling distribution of the mean of the number of calories, for samples of size 3.
		
```{r}

```



*	Create a histogram and describe the simulated sampling distribution.
		
```{r}

```


Description of sampling distribution:		
		
*	Use the simulation to find the probability that the average calorie content of 3 beers is less than 150 calories.
		
```{r}
```
		
* Use the simulation to find the probability that the average calorie content of 3 beers is more than 160 calories.
  
```{r}
```
  
  
4. The following code generates two ecdfs, one for calories of ales, one for calories of lagers. 
 
```{r}
Ales <- subset(BeerData, select=Calories, subset=Type=="Ale", drop=T)
Lagers <- subset(BeerData, select=Calories, subset=Type=="Lager", drop=T)
#plot.new()
plot.ecdf(Ales, col="red")
plot.ecdf(Lagers, col="blue", add=TRUE)
abline(v=165)
```

What percentage of ales in the data set have less than 165 calories? 

What percentage of lagers in the data set have less than 165 calories?


# Permutation test

Conduct a permutation test to determine whether there is a difference in the mean calorie content between ales and lagers. Your test should start with a statement of the formal hypotheses and finish with a conclusion.

Put each step in its own chunk and **clearly explain what each chunk is doing and what information the output provides.**

```{r}

```




```{r}

```



```{r}

```


```{r}

```



```{r}

```


(I put in several r-chunks.  You might use more, you might use fewer.)


# Critical reading

Below, I conduct a permutation test to test whether there was, on average, a difference in the annual snow fall in Colorado in the years 2019 and 2020.  Identify the mistake and fix it.

First, we load and briefly inspect the data:
```{r}
SnowData <- read.csv("https://www.dropbox.com/s/h8lkjdqoqui3bzw/TestDownload2.csv?dl=1")
head(SnowData)
```

The data shows annual snow fall recorded at 51 NOAA weather station in Colorado in 2019, and 2020. 


The formal statements of the hypotheses are:

$$H_0: \mu_{d_{2020}-d_{2019}}=0$$

$$H_A: \mu_{d_{2020}-d_{2019}}\neq 0$$

where $d_{2020}-d_{2019}$ represents the difference in snow fall by station. 

We'll compute the difference in snowfall at each station and compute the observed mean of differences:

```{r}
DiffAtStations<-SnowData$Snow2020-SnowData$Snow2019
Observed<-mean(DiffAtStations)
Observed
```

Next, we'll pool the data in order to find a permutation distribution:

```{r}
PooledData<- c(SnowData$Snow2020, SnowData$Snow2019)
```

Here is the loop that generates a permutation distribution:


```{r}
N=10^4
SampleDist<-numeric(N)
for (i in 1:N)
{index<-sample(102,51, replace=F)
SampleDist[i]<- mean(PooledData[index]-PooledData[-index])}
```



Finally, we compute the $p$-value for our observed statistic:

```{r}
2*min((sum(SampleDist >= Observed)+1)/(N+1),
(sum(SampleDist <= Observed)+1)/(N+1))
```

Since the $p$-value is small we reject the null hypothesis and conclude that the mean difference in snow depth between 2019 and 2020 was different.


**Your notes are below.  Do NOT delete the following chunk or your file may not knit.
```{r}
knitr::knit_exit()
```


Chapter 1

Observation: Each row of the data set. A value of something of interest you're measuring or counting during a study or experiment: a person's height, a blank account value at a certain point in time, or number of animals.

Variable: Each column. An attribute that describes a person, place, thing, or idea: a person's hair color is a potential variable, which could have the calue of "blond" for one person and "brunette" for another.

Population/sample: Represents all the individual cases, whether they are babies, fish, cars, or coin flips.

Random Sample: a sample of independent and identically distributed (i.i.d.) observations from the population, if the population is infinite.

Numeric Variable: A variable where the measurement or number has a numerical meaning: total rainfall measured in inches, heart rate, number of cheeseburgers consumed in an hour.

Factor Variable: The variables that experimenters control during an experiment in order to determine their effect on the response variable.

Random Variable: A variable whose value is unkown or a function that assigns values to each of an experiment's outcomes.

Independent and Identically distributed:If each random variable has the same probability distribution as the others and all are mutually independent.

Parameter: A (numerical) characteristic of a population of of a probability distribution

Statistic: A (numerical) characteristic of data.

Statistical Inference: Drawing a conclusion about a population based on information about a sample.

Census: Take a poll of every individual in the population.

Sampling Frame: A list from which the researchers will choose their sample.

Under coverage: Omitting people from the target population in the survey.

Probability (scientific) Sampling Scheme: A scheme that gives everybody in the population a positive chance of being selected.

Stratified Sampling: The population is divided into non overlapping strata, and then random samples are drawn from each stratum.

Cluster Sampling: The population is divided into non overlapping clusters, and then a random sample of clusters is drawn.

Observational Study: A study in which researchers observe participants but do not influence the outcome. No cause and effect relationship. Researchers CANNOT conclude anything from this study because it was only observational.

Experiment: Researchers will manipulate the environment in some way to observe the response of the objects of interest (people, mice, ball bearings, etc.).

Subjects: When the objects of interest in an experiment are people.

Experimental Units: When the objects of interest in an experiment are not people.

Experimental Studies: Are ones where researchers introduce an intervention and study the effects.

Random Assignment: To spread out the effects of unknown or uncontrollable factors that might introduce unwanted variability into the results.

Double-blind Experiment: One in which neither the researcher nor the subject knows who is receiving which treatment.

Single-blinded Experiment: Just the researcher or the subject (but not both) know who is receiving which treatment.

Bias: The systematic favoring of one outcome over another.

Sampling with replacement: A sampling unit is drawn from a finite population and is returned to that population, after its characteristic(s) have been recorded.


---------------------------------------------------
Factors: how many columns there are (along the top)
Numeric: which ones have number values
Factor: which ones have words for the inputs
Number of observations: how many entries there are

Import File:
   Name it (Girls) whatever you want.
   Change the end: Girls2004 to the data set you want
```{r}
Girls <- read.csv("https://sites.google.com/site/chiharahesterberg/data2/Girls2004.csv")
Girls2 <- read.csv("Data/Girls2004.csv")
```

To see data:
```{r}
#head(Girls)
```

-----------------------------------------------------
Suppose we want to look at the birth weight of girls born in Wyoming.  We can extract this information from the data set with the "subset" command:

```{r}
WYGirls <- subset(Girls, select=Weight, subset=State=="WY", drop=T)
WYGirls
```
--------------------------------------------------------
Want to take the mean:
  The second will trim the numbers 25% on each side
```{r}
mean(WYGirls)
mean(WYGirls, trim=.25)
```


Chapter 2

Barplot and Table:

```{r}
FlightData <-  read.csv("https://sites.google.com/site/chiharahesterberg/data2/FlightDelays.csv")
```

```{r}
barplot(table(FlightData$Carrier))
table(FlightData$Carrier)
```
--------------------------------------------------
Contingency tables:
** In this case we want the second proptable
  From the table, add AA yess and nos and divide no by the sum of them.
  
```{r}
table(FlightData$Carrier, FlightData$Delayed30)
prop.table(table(FlightData$Carrier, FlightData$Delayed30))
prop.table(table(FlightData$Carrier, FlightData$Delayed30),1)
prop.table(table(FlightData$Carrier, FlightData$Delayed30),2)
```
----------------------------------------------------------------------
Histogram:
  This histogram is skewed right
```{r}
hist(FlightData$Delay)
```

THis is only for UA flight delays:
```{r}
UAOnly <- subset(FlightData, subset=Carrier=="UA")
dim(UAOnly)
hist(UAOnly$Delay)
```

To cut the sides down on a histogram:
```{r}
width=seq(-50,700, by=25)
hist(UAOnly$Delay, width)
```
----------------------------------------------------------------------------
Take the median, min, max, and range
```{r}
median(UAOnly$Delay)
max(UAOnly$Delay)
min(UAOnly$Delay)
range(UAOnly$Delay)
```
Show the percentages and for the given quantiles (IQR)
```{r}
quantile(UAOnly$Delay)
```
may need to add the end
  quantile(Recidivism$Days,na.rm = TRUE)

Standard deviation:
```{r}
sd(UAOnly$Delay)
```

---------------------------------------------------------------------------
To find the numeric summaries of delay and carrier
```{r}
tapply(FlightData$Delay, FlightData$Carrier, summary)
tapply(FlightData$Delay, FlightData$Carrier, mean)
tapply(FlightData$Delay, FlightData$Day, sd)
```
---------------------------------------------------------------------------
Boxplot:

```{r}
boxplot(FlightData$Delay~FlightData$Destination)
```

```{r}
remove <- subset(FlightData, subset=Delay<=100)
boxplot(remove$Delay~remove$Destination)
```
-----------------------------------------------------------------------------
Quantile-Quantile plots (qq plots):
  This plot is skewed right
  If the Normal Q-Q Plot line follows the dots, the plot is approximately a normal             distribution.
```{r}
plot.new()
qqnorm(FlightData$Delay)
qqline(FlightData$Delay)
```
-------------------------------------------------------------------------------
Summary of data:
```{r}
summary(FlightData$Delay)
```
-------------------------------------------------------------------------------
Scatter plot:
```{r}
plot(FlightData$Carrier, FlightData$Delay, main = 'Scatterplot')
```
-----------------------------------------------------------------------------
ecdf: of ages over and under 25 that were in prison.
```{r}
Under25 <- subset(Recidivism, select=Days, subset=Age25=="Under 25", drop=T)
head(Under25)
Over25 <- subset(Recidivism, select=Days, subset=Age25=="Over 25", drop=T)
head(Over25)
```

```{r}
plot.ecdf(Under25, col="red")
plot.ecdf(Over25, col="blue", add=TRUE)
abline(v=400)
```
------------------------------------------------------------------------------

Chapter 3


Permutation Test: (1) Analyze the problem (2) Choose a test statistic (3) Resample and recompute the test statistic (4) Reject or fail to reject the null hypothesis
---
Difference in means:

* `N` is the number of times we'll resample
* `result` is a vector for storing the results of each resample
* the loop is each iteration of the resample

N <- 10^4-1
result <- numeric(N)
for (i in 1:N)
{index <- sample(4029,size=1123, replace=F)
result[i] <- mean(alldelays[index])-mean(alldelays[-index])
} 
-----
For delayed more than 20 min:

poolsize <- length(alldelays)
GroupOneSize <- length(Maydelays)
GroupTwoSize <- length(Junedelays)

N <- 10^4-1
result <- numeric(N)
for (i in 1:N)
{index <- sample(poolsize,size=GroupOneSize, replace=F)
result[i] <- (sum(alldelays[index]>20)/GroupOneSize)-(sum(alldelays[-index]>20)/GroupTwoSize)

hist(result)


------------------------------------------------------------------------------------
Chapter 4

Sampling Distribution: A probability distribution of a statistic obtained from a larger number of samples drawn from a specific population. The sampling distribution of a given population is the distribution of frequencies of a range of different outcomes that could possibly occur for a statistic of a population.   



 Example:
A and B are two populations:
```{r}
A <- c(3, 5, 7, 9, 10, 16)
B <- c(8, 10, 11, 15, 18, 25, 28)
max(A)
max(B)
```

a) Draw random samples without replacement of size 3 from each population, and simulate the sampling distribution of the sum of their max.

```{r}
N <- 10^4
Xa <- numeric(N)
for (i in 1:N)
{
  sampA <- sample(A, 3, replace = FALSE)
  sampB <- sample(B, 3, replace = FALSE)
  Xa[i] <- max(sampA)+max(sampB)
}

stripchart(A,method='stack',at=0)
stripchart(B,method='stack',at=0)
hist(Xa)
```

b) Use your simulation to estimate the probability  that maximum of the union is leass than 20.

```{r}
prob2 <- mean(Xa<20)
length(which(Xa<20))/N
```
Therefore, the probability that the sum of the maximums is less than 20 is 0.0022.

c)Draw random samples of size 3 from A and B, find the max of union of the two sets. Simulate the sampling distribution of that. Commpare with part a 

```{r}
N <- 10^4
Xb <- numeric(N)
for (i in 1:N)
{
  sampA <- sample(A, 3, replace = FALSE)
  sampB <- sample(B, 3, replace = FALSE)
  Xb[i] <- max(union(sampA,sampB))
}

hist(Xb)
```

d)Find the probability tht the max of the union is less than 20

```{r}
prob2 <- mean(Xb<20)
length(which(Xb<20))/N
```
Therefore, the probability that the sum of the maximums is less than 20 is 0.2854.



-----------------------------------------------------
```{r}
HSD <- 0.286
```


```{r}
Prob <- pbinom(230,800,HSD)-pbinom(219,800,HSD)
ans <- 'The probablility that between 220 and 230 US adults have a high school degree is'
cat(ans,Prob)
```

```{r}
plot(dbinom(0:800,800,HSD), type = 'b')
```

## Another Problem

```{r}
role6 <- 1-(5/6)^6
```


```{r}
P <- pbinom(1:6,6,role6)
ans <- 'The probablility that between 220 and 230 US adults have a high school degree is'
cat(ans,P)
```








