---
title: "Exam_One_A"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





# Terminology
1. What is the difference between a *population* and a *sample*?

Answer: 


2. What is the difference between a *statistic* and a *parameter*.  Give an example of a *statistic*.

Answer:


3. In a data set, what is the difference between an *observation* and a *variable*?

Answer:



# Exploratory Data Analysis

Use one of the following two chunks to load the data set `Alelager`.

If you have the `resampledata` package installed use:

```{r}
library(resampledata)
BeerData<-Alelager
head(BeerData)
```

**or**, you can use:

```{r}
BeerData<-read.csv("https://sites.google.com/site/chiharahesterberg/data2/Alelager.csv")
head(BeerData)
```

1. Create side-by-side box plots of the alcohol content, grouped by beer type. Comment on the information the box plots convey.

```{r}

```

Commentary: 


2. Create a histogram of the calorie content of all beers in the data set. Describe the distribution.
```{r}

```

Description of distribution:

3.  Sampling Distribution
*	Simulate the sampling distribution of the mean of the number of calories, for samples of size 3.
		
```{r}

```



*	Create a histogram and describe the simulated sampling distribution.
		
```{r}

```


Description of sampling distribution:		
		
*	Use the simulation to find the probability that the average calorie content of 3 beers is less than 150 calories.
		
```{r}
```
		
* Use the simulation to find the probability that the average calorie content of 3 beers is more than 160 calories.
  
```{r}
```
  
  
4. The following code generates two ecdfs, one for calories of ales, one for calories of lagers. 
 
```{r}
Ales <- subset(BeerData, select=Calories, subset=Type=="Ale", drop=T)
Lagers <- subset(BeerData, select=Calories, subset=Type=="Lager", drop=T)
#plot.new()
plot.ecdf(Ales, col="red")
plot.ecdf(Lagers, col="blue", add=TRUE)
abline(v=165)
```

What percentage of ales in the data set have less than 165 calories? 

What percentage of lagers in the data set have less than 165 calories?


# Permutation test

Conduct a permutation test to determine whether there is a difference in the mean calorie content between ales and lagers. Your test should start with a statement of the formal hypotheses and finish with a conclusion.

Put each step in its own chunk and **clearly explain what each chunk is doing and what information the output provides.**

```{r}

```




```{r}

```



```{r}

```


```{r}

```



```{r}

```


(I put in several r-chunks.  You might use more, you might use fewer.)


# Critical reading

Below, I conduct a permutation test to test whether there was, on average, a difference in the annual snow fall in Colorado in the years 2019 and 2020.  Identify the mistake and fix it.

First, we load and briefly inspect the data:
```{r}
SnowData <- read.csv("https://www.dropbox.com/s/h8lkjdqoqui3bzw/TestDownload2.csv?dl=1")
head(SnowData)
```

The data shows annual snow fall recorded at 51 NOAA weather station in Colorado in 2019, and 2020. 


The formal statements of the hypotheses are:

$$H_0: \mu_{d_{2020}-d_{2019}}=0$$

$$H_A: \mu_{d_{2020}-d_{2019}}\neq 0$$

where $d_{2020}-d_{2019}$ represents the difference in snow fall by station. 

We'll compute the difference in snowfall at each station and compute the observed mean of differences:

```{r}
DiffAtStations<-SnowData$Snow2020-SnowData$Snow2019
Observed<-mean(DiffAtStations)
Observed
```

Next, we'll pool the data in order to find a permutation distribution:

```{r}
PooledData<- c(SnowData$Snow2020, SnowData$Snow2019)
```

Here is the loop that generates a permutation distribution:


```{r}
N=10^4
SampleDist<-numeric(N)
for (i in 1:N)
{index<-sample(102,51, replace=F)
SampleDist[i]<- mean(PooledData[index]-PooledData[-index])}
```



Finally, we compute the $p$-value for our observed statistic:

```{r}
2*min((sum(SampleDist >= Observed)+1)/(N+1),
(sum(SampleDist <= Observed)+1)/(N+1))
```

Since the $p$-value is small we reject the null hypothesis and conclude that the mean difference in snow depth between 2019 and 2020 was different.



**Your notes are below.  Do NOT delete the following chunk or your file may not knit.
```{r}
knitr::knit_exit()
```



Terminology:

Observation: The rows of the data sets. The thing or number that has been observed.
Variable: The columns of a data set, some characteristic that is obtained from each observation.
Population: All the individual cases.
(Random) Sample: A sample of independent and identically distributed observations from the population.
Numeric Variable: A data variable that take on any value within a finite of infinite interval. 
Factor Variable: Categorical variable that can be either numeric or string variables.
Random variable:a variable whose value is unknown or a function that assigns values to each of an experiment's outcomes.
Independent and Identically Distributed:
Sampling with Replacement:Draw an observation from the population and then placing that observation back in the population.
Sampling without Replacement: Drawing an observation for a population and not placing it back in the population. 
Parameter: A numerical characteristic of a population of of a probability distribution.
Statistic: A numerical characteristic of data. 
Statistical Inference: Drawing a conclusion about a population based on information about a sample.
Observations Study: A study in which researchers observe participants but do not influence the outcome.
Experiment(al Study):Researcher manipulate the environment in some way to observe the respince of the objects of interest.  


Uploading Data:
```{r}
#Name <- read.csv("https://sites.google.com/site/chiharahesterberg/data2/FlightDelays.csv")
  #Will have to change the name of the data set that is being read the to the CSV file
```

Basic Commands:
```{r}
#table(Name)
  #Creates a table of the data set
#table(Name$Delays)
  #Creates a table of the specified set of data from the data set
#prop.table(table(FlightData$Day, FlightData$Delayed30),1)
  #Creates a proportion table of the specified data. 1 computes rows, 2 computes columns
#barplot(table(name))
  #Creates a bar plot from the specified table
#boxplot(FlightData$FlightLength~FlightData$Delayed30)
  #Creates a box plot of the specified data
#hist(Name)
#width = seq(0,20, by=1)
  # How to set teh width on a histogram
  #Creates a histogram of the specified data

```

Creating Subsets:
```{r}
#Smoker <- subset(BirthsData, select=Weight, subset=Smoker=="Yes", drop=T)
  #Can specify what type of data is included in by manipulating the names
```

ECDF Plot (Empirical Cumulative Distrabution Function):
```{r}
#plot.new()
#plot.ecdf(Smoker, col="red")
#plot.ecdf(nonsmoker, col="blue", add=TRUE)
#abline(v=3000)
```

Permutation Testing:
```{r}
#delays <- FlightDelays
#head(delays)

#UAdelays <- subset(delays, select=Delay, subset=Carrier=="UA", drop=T)

#AAdelays <- subset(delays, select=Delay, subset=Carrier=="AA", drop=T)

#observed <- mean(UAdelays)-mean(AAdelays)
#observed
  #Create subsets from the original data that are going to be compaired

#length(UAdelays)
#length(AAdelays)
#length(UAdelays)+length(AAdelays)
  # Check to make sure that the numbers make sense.

#AllDelays <- FlightDelays$Delay
#poolsize <- length(AllDelays)
#poolsize

#GroupOneSize <- length(UAD)
#GroupOneSize

#GroupTwoSize <- length(AAD)
#GroupTwoSize
  #Set the group sizes for the resampling

#N <- 10^6-1
#Result <- numeric(N)
#for (i in 1:N)
#{index <- sample(poolsize, size=GroupOneSize, replace=F)
#Result[i] <- (sum(AllDelays[index]>20)/GroupOneSize)-(sum(AllDelays[-index]>20)/GroupTwoSize)}
  # Run the re-sampling loop

#hist(Result)
#2*min((sum(Result >= Observed)+1)/(N+1),sum(Result <= Observed)+1/(N+1))
  #Check the results and find the P-value
  #Less than 0.05 is statically significant results
```

Matched Pairs:
```{r}
#DiffInPairs <- Groceries$Target-Groceries$Walmart
#DiffInPairs
  #Find the differance in pairs

#Observed <- mean(DiffInPairs)
#Observed
  #Find the observation

#summary(DiffInPairs)
#sd(DiffInPairs)
  #find the statics summery for the data

#Poolsize <- length(Groceries$Product)
#Poolsize
  #Find the pool size

#sample(c(-1,1),Poolsize, replace=T)
  #Sample the data

#N <- 10^6-1
#Result <-numeric(N)
#for (i in 1:N)
#{Sign <- sample(c(-1,1),Poolsize, replace=T)
  #Diff <- Sign*DiffInPairs
#Result[i] <- mean(Diff)
#}
  #Run the test for the data

#hist(Result)
#2*min((sum(Result >= Observed)+1)/(N+1),(sum(Result <= Observed)+1)/(N+1))
  #Check the results and calculate the P-value
```

Distributions:
```{r}
#N <- 10^5
#Xbar <- numeric(N)
#for(i in 1:N)
#{Sample <- sample(Pop,4,replace=TRUE)
 #Xbar[i] <- mean(Sample)}
    #The loop for creating a smapleing distrabution

#HSD1 <- pbinom(220,800,HSD)
  #Use one less than given number so not to cut one out.

#barplot(dbinom(0:800, 800, success))
  #For plotting the distrabution
```

Other Distributions:
```{r}
#Geometric Distrabution
  #dgeom(3, basketsuccess)
  #barplot(dgeom(0:20, basketsuccess),names)

#Exponential
  #pexp(1/2, 3)
  #qexp(.9,3)

#Uniform distribution
  #library(ggfortify)
  #ggdistribution(dunif, seq(-3, 23, 0.1), min=2, max=13)

  #punif(10, 2, 13)
  #qunif(.7, 2, 13)

```







