---
title: "Exercise_5_9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Consider a population that has gamma distribution with parameters $r=5$ and $\lambda=\frac{1}{4}$:

```{r}
library(ggfortify)
r <- 5
lambda <- 1/4
ggdistribution(dgamma, seq(0, 80, 0.01), shape=r, rate=lambda)
```

a) First we generate a *sampling distribution*:

```{r}
N<-10^4
n<-200
sampdist<- numeric(N)
for (i in 1:N)
{samp<- rgamma(n, shape=r, rate=lambda)
sampdist[i]<-mean(samp)}
```

Here is the plot:

```{r}
hist(sampdist)
```

We find the mean and standard deviation of the sampling distribution:

```{r}
mean(sampdist)
sd(sampdist)
```


## Bootstrap process:

b) First draw a base sample:

```{r}
BaseSamp<-rgamma(n, shape=r, rate=lambda)
```

Its distribution:

```{r}
hist(BaseSamp)
```

which is skewed right.

The mean and standard deviation of this *one* sample is:

```{r}
mean(BaseSamp)
sd(BaseSamp)
```

c) Now we bootstrap:

```{r}
N=1000
BootDist <- numeric(N)
for (i in 1:N)
{samp <- sample(BaseSamp, n, replace=T)
BootDist[i] <- mean(samp)}
```

Here is the plot of the bootstrap distribution:

```{r}
hist(BootDist)
```

Here are the mean and standard deviation of the bootstrap distribution:

```{r}
mean(BootDist)
sd(BootDist)
```
Here are the mean and standard deviation of the sample distribution:

```{r}
mean(sampdist)
sd(sampdist)
```




```{r}
Distribution <- c("Population", "Sample Distribution", "Sample", "Bootstrap Distribution")
means <- c(r/lambda, mean(sampdist), mean(BaseSamp), mean(BootDist))
sdevs <- c(sqrt(r)/lambda, sd(sampdist), sd(BaseSamp),sd(BootDist))
results <- cbind(Distribution, means, sdevs)
knitr::kable(results, digits=2)
```

