---
title: "OtherDistributionsHW"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1.
---
As you walk down the hall, the probability that the next person you see is wearing some sort of Skyhawk apparel is 0.15. 

**a)** 
What is the probability that you walk past exactly 20 people before seeing one with Skyhawk apparel?

```{r}
SkyhawkSuccess <- 0.15
dgeom(20, SkyhawkSuccess)
```

**b)** 
What is the probability that you walk past at least 20 people before seeing one with Skyhawk apparel?

```{r}
1-pgeom(20, SkyhawkSuccess)
```

**c)** 
What is the probability that 4 of the next 20 people you walk past are wearing Skyhawk apparel?

```{r}
dbinom(4,20, SkyhawkSuccess)
```

2)
-----
It snows once every 7 days during the Winter. 

**a)**
What is the probability that it will snow sometime in the next 3 days?

```{r}
pexp(3, 1/7)
```

**b)**
What is the probability that it will snow next Thursday at 3pm exactly?

zero

**c)**
For how long should I plan my vacation if I want to be 80% certain of it snowing?




4.16
----
**a)** 10

**b)**

```{r}
N=1000
SampDist <- numeric(N)
for (i in 1:N)
{SampFromExp <- rexp(30, 1/10)
SampDist[i] <- mean(SampFromExp)}

hist(SampDist)
mean(SampDist>=12)
```

**c)**

No.




