---
title: "Permutation_Test_2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load the BookPrices data set.  Un-comment ONE of the following chunks to do this:

```{r}
#library(resampledata)
#Books <- BookPrices
#head(Books)
```

```{r}
#Books<-read.csv("https://sites.google.com/site/chiharahesterberg/data2/BookPrices.csv")
#head(Books)
```

Conduct a permutation hypothesis test to determine if there is a difference in the mean price of Math books and the mean price of Economics books. Your test should start with a statement of the formal hypotheses and finish with a conclusion.


You may need to pool the prices of math books and the prices of physics books. Here is an example of pooling data from two vectors:

```{r}
odds <- c(1,3,5,7)
evens <- c(0,2,4,6,8,10)
pool <- c(odds, evens)
pool
```

If you want to see all the subject areas in the data set use the following chunk.

```{r}
levels(Books$Subject)
```

Here is the piece of code I always seem to forget part of:

```{r}
2*min((sum(result >= observed)+1)/(N+1),
(sum(result <= observed)+1)/(N+1))
```

I recommend restarting R since we tend to reuse the same variable names from one exercise to another.


